---
title: "Data Analysis Plan RR JCL"
author: "Catherine Laing"
date: "09/01/2020"
output: html_document
---

```{r setup, include=FALSE}
source("dummy data RR.R") # run simulated data to generate sample_data_P1P2 and sample_data_P3
```

```{r add libraries}
library(ggplot2)
library(lme4)
library(readr)
library(forcats)
library(effsize)
library(rcompanion)

```

```{r check for outliers}

# Outliers defined as any productions that vary more than 3 standard deviations from the mean in each direction
# Group by experiments as we are expecting differences across paradigms

sample_data_P1P2 %>% group_by(paradigm) %>%
  mutate(outlier = (MOTpitch > 
                      (mean(MOTpitch, na.rm=T) + 
                         3*(sd(MOTpitch, na.rm=T))) |
                      MOTpitch <
                      (mean(MOTpitch, na.rm=T) - 
                         3*(sd(MOTpitch, na.rm=T))))) %>%
  filter(outlier == TRUE)  

sample_data_P3 %>% group_by(paradigm) %>%
  mutate(outlier = (MOTpitch > 
                      (mean(MOTpitch, na.rm=T) + 
                         3*(sd(MOTpitch, na.rm=T))) |
                      MOTpitch <
                      (mean(MOTpitch, na.rm=T) - 
                         3*(sd(MOTpitch, na.rm=T))))) %>%
  filter(outlier == TRUE)  

# remove any outliers from analysis and re-run to test for differences in alpha level with and without outliers
```

# Descriptive statistics


```{r descriptive statistics between paradigms}

sample_data_P1P2 %>%                                     
  summarise(meanpitch = mean(MOTpitch, na.rm=T),
            sdpitch = sd(MOTpitch, na.rm = T))

sample_data_P1P2 %>% group_by(paradigm) %>%               # mean across conditions, P1 & P2         
  summarise(meanpitch = mean(MOTpitch, na.rm=T),
            sdpitch = sd(MOTpitch, na.rm = T))

sample_data_P3 %>%                                       # mean across conditions   , P3            
  summarise(meanpitch = mean(MOTpitch, na.rm=T),
            sdpitch = sd(MOTpitch, na.rm = T))

```



# Analysis of potentially confounding variables

```{r checking for confounding variables}

# Does infant age or sex, or experimental item, affect mothers' pitch?

# Age

age.model.P1P2 <- lm(MOTpitch ~ age, data = sample_data_P1P2)
summary(age.model.P1P2)

age.model.P3 <- lm(MOTpitch ~ age, data = sample_data_P3)
summary(age.model.P3)

# Sex

sex.model.P1P2 <- lm(MOTpitch ~ sex, data = sample_data_P1P2)
summary(sex.model.P1P2)

sex.model.P3 <- lm(MOTpitch ~ sex, data = sample_data_P3)
summary(sex.model.P3)

# item

item.model.P1P2 <- lm(MOTpitch ~ item, data = sample_data_P1P2)
summary(item.model.P1P2)

item.model.P3 <- lm(MOTpitch ~ item, data = sample_data_P3)
summary(item.model.P3)

```

# Analysis 1: Paradigms 1 & 2

```{r paradigms 1 & 2 descriptive statistics}

# Size

sample_data_P1P2 %>%                                         # Across paradigms
  group_by(size) %>%
  summarise(meansizeP1P2 = mean(MOTpitch, na.rm=T),
            sdsizeP1P2 = sd(MOTpitch, na.rm=T))  

sample_data_P1P2 %>%                                         # Paradigms  1 and 2, within paradigms
  group_by(size, paradigm) %>%
  summarise(meansizeP1P2 = mean(MOTpitch, na.rm=T),
            sdsizeP1P2 = sd(MOTpitch, na.rm=T))  


# Congruence

sample_data_P1P2 %>%                                         # Across paradigms
  group_by(cong) %>%
  summarise(meancongP1P2 = mean(MOTpitch, na.rm=T),
            sdcongP1P2 = sd(MOTpitch, na.rm=T))  

sample_data_P1P2 %>%                                         # Paradigms  1 and 2, within paradigms
  group_by(cong, paradigm) %>%
  summarise(meancongP1P2 = mean(MOTpitch, na.rm=T),
            sdcongP1P2 = sd(MOTpitch, na.rm=T))    



```

# Figure 1A-B: Plotting mothers' mean pitch across size and expected size variables in free play setting vs. semi-structured play

```{r Figure 1A}

paradigm_numbers <- c("-1" = "Paradigm 1",
                      "1" = "Paradigm 2",
                      "3" = "Paradigm 3")

size <- c("-1" = "Large objects",
          "1" = "Small objects")

Figure1A <- ggplot(subset(sample_data_P1P2), aes(x = as.factor(size), y = MOTpitch, color = as.factor(size))) +
  geom_point(aes(group=size), shape = 1, size = 4, position=position_dodge(-.2)) +
  stat_summary(geom = "pointrange", 
               size=1.5,  
               colour = "red") +
    scale_x_discrete(limits = c("1", "-1"),
                   labels = c("small", "large")) +
  xlab("Actual Object Size") +
  ylab("Mothers' pitch (Hz)") +
  ggtitle("Size difference") +
  theme_bw(base_size=25) +
   theme(axis.text = element_text(size = 20),
         axis.title = element_text(size=20),
         legend.position = "none") +
  facet_wrap(~paradigm, ncol = 2, labeller = as_labeller(paradigm_numbers))
plot(Figure1A)
```

```{r Figure 1B}

Figure1B <- ggplot(subset(sample_data_P1P2, paradigm == -1), aes(x = as.factor(cong), y = MOTpitch)) +
  geom_point(aes(group=as.factor(cong), color = as.factor(cong)), shape = 1, size = 4, position=position_dodge(-.2)) +
    stat_summary(geom = "pointrange", 
               size=1.5,  
               colour = "red") +
  scale_x_discrete(limits = c("1", "-1"),
                   labels = c("congruent", "reversed")) +
  xlab("Real-world congruence") +
  ylab("Mothers' pitch (Hz)") +
  ggtitle("Expected Object Size: Paradigm 1") +
  theme_bw(base_size=25) +
   theme(axis.text = element_text(size = 20),
         axis.title = element_text(size=20),
         legend.position = "none")  +
  facet_wrap(~size, ncol = 2, labeller = as_labeller(size))

plot(Figure1B)
```

```{r paradigm 1 & 2 modelling differences across variables}

modelP1P2 <- lmer(MOTpitch ~ size*cong*paradigm + 
                           (1+size*cong*paradigm|subj) + 
                           (1+size*cong*paradigm|item) + 
                           (1+size*cong*paradigm|word), 
                            data = sample_data_P1P2,  REML=FALSE) 

modelP1P2.null.size <- lmer(MOTpitch ~ cong*paradigm + 
                           (1+size*cong*paradigm|subj) + 
                           (1+size*cong*paradigm|item) + 
                           (1+size*cong*paradigm|word), 
                            data = sample_data_P1P2, REML=FALSE)

modelP1P2.null.cong <- lmer(MOTpitch ~ size*paradigm + 
                           (1+size*cong*paradigm|subj) + 
                           (1+size*cong*paradigm|item) + 
                           (1+size*cong*paradigm|word), 
                            data = sample_data_P1P2, REML=FALSE)

modelP1P2.null.paradigm <- lmer(MOTpitch ~ size*cong + 
                           (1+size*cong*paradigm|subj) + 
                           (1+size*cong*paradigm|item) + 
                           (1+size*cong*paradigm|word), 
                            data = sample_data_P1P2, REML=FALSE)

anova(modelP1P2, modelP1P2.null.size) # test for effect of object size on mothers' pitch
anova(modelP1P2, modelP1P2.null.cong) # test for effect of object congruence on mothers' pitch
anova(modelP1P2, modelP1P2.null.paradigm) # test for effect of paradigm on mothers' pitch

## Shayne: Need to add other comparisons here, I think

```


# Analysis 2: Paradigm 3

```{r paradigm 3 descriptive statistics}

# Size

sample_data_P3 %>% 
  group_by(size) %>%
  summarise(meansizeP3 = mean(MOTpitch, na.rm=T),
            sdsizeP3 = sd(MOTpitch, na.rm=T))  

# Colour (control condition)

sample_data_P3 %>% 
  group_by(colour) %>%
  summarise(meancolourP3 = mean(MOTpitch, na.rm=T),
            sdscolourP3 = sd(MOTpitch, na.rm=T))   
 

```

# Figure 2A&B: Plotting mothers' mean pitch according to size (test) and colour (control) of object in Paradigm 3

```{r Figure 2A}

Figure2A <- ggplot(sample_data_P3, aes(x = as.factor(size), y = MOTpitch)) +
  geom_point(aes(group=as.factor(size), color = as.factor(size)), shape = 1, size = 4, position=position_dodge(-.2)) +
    stat_summary(geom = "pointrange", 
               size=1.5,  
               colour = "red") +
  scale_x_discrete(limits = c("1", "-1"),
                   labels = c("Large", "Small")) +
  xlab("Object size") +
  ylab("Mothers' pitch (Hz)") +
  ggtitle("Object Size, Paradigm 3") +
  theme_bw(base_size=25) +
   theme(axis.text = element_text(size = 20),
         axis.title = element_text(size=20),
         legend.position = "none")

plot(Figure2A)
```

```{r Figure 2B}

Figure2B <- ggplot(sample_data_P3, aes(x = as.factor(colour), y = MOTpitch)) +
  geom_point(aes(group=as.factor(colour), color = as.factor(colour)), shape = 1, size = 4, position=position_dodge(-.2)) +
    stat_summary(geom = "pointrange", 
               size=1.5,  
               colour = "red") +
  scale_x_discrete(limits = c("1", "-1"),
                   labels = c("ColourA", "ColourB")) +
  xlab("Object Colour") +
  ylab("Mothers' pitch (Hz)") +
  ggtitle("Object Colour, Paradigm 3") +
  theme_bw(base_size=25) +
   theme(axis.text = element_text(size = 20),
         axis.title = element_text(size=20),
         legend.position = "none")

plot(Figure2B)
```



```{r modelling differences across variables in paradigm 3}

modelP3 <- lmer(MOTpitch ~ size + colour + (1+size+colour|subj) + (1+size+colour|item), data = sample_data_P3, REML = FALSE)

modelP3.null.size <- lmer(MOTpitch ~ colour + (1+size+colour|subj) + (1+size+colour|item), data = sample_data_P3, REML = FALSE)

modelP3.null.colour <- lmer(MOTpitch ~ size + (1+size+colour|subj) + (1+size+colour|item), data = sample_data_P3, REML = FALSE)

anova(modelP3, modelP3.null.size) # test for effect of size
anova(modelP3, modelP3.null.colour) # test for effect of colour


# Shayne: we also need to confirm this model and its comparisons

```

# Analysis 3: Differences across paradigms

```{r cross-paradigm descriptives}

sample_data_P123 %>% 
  group_by(paradigm) %>%
  summarise(meanpitch = mean(MOTpitch, na.rm=T),
            sdspitch = sd(MOTpitch, na.rm=T))   
 
```

```{r testing for pitch~size differences with paradigm as a factor}

modelP123 <- lmer(MOTpitch ~ size*paradigm + 
                    (1+size*paradigm|subj) + 
                    (1+size*paradigm|word), 
                  data = sample_data_P123, REML = FALSE)

modelP123.null.size <- lmer(MOTpitch ~ paradigm + 
                    (1+size*paradigm|subj) + 
                    (1+size*paradigm|word), 
                  data = sample_data_P123, REML = FALSE)

modelP123.null.paradigm <- lmer(MOTpitch ~ size + 
                    (1+size*paradigm|subj) + 
                    (1+size*paradigm|word), 
                  data = sample_data_P123, REML = FALSE)

anova(modelP123, modelP123.null.size)
anova(modelP123, modelP123.null.paradigm)

```


```{r Figure 3}

Figure3 <- ggplot(subset(sample_data_P123), aes(x = as.factor(size), y = MOTpitch, color = as.factor(size))) +
  geom_point(aes(group=size), shape = 1, size = 4, position=position_dodge(-.2)) +
  stat_summary(geom = "pointrange", 
               size=1.5,  
               colour = "red") +
    scale_x_discrete(limits = c("1", "-1"),
                   labels = c("small", "large")) +
  xlab("Actual Object Size") +
  ylab("Mothers' pitch (Hz)") +
  ggtitle("Size difference") +
  theme_bw(base_size=25) +
   theme(axis.text = element_text(size = 20),
         axis.title = element_text(size=20),
         legend.position = "none") +
  facet_wrap(~paradigm, ncol = 3, labeller = as_labeller(paradigm_numbers))

plot(Figure3)

```
